-----
title: "`r subdir_name`"
author: "Kadi Vaher"
date: "`r Sys.time()`"
output: html_document
---

```{r setup, include=T, message=F}
subdir_name <- "microbiome_demotable"

# load packages
library(tidyverse)
library(here)
library(magrittr)
library(glue)
library(ggpubr)
library(knitr)
library(MVN)
library(tableone)

# set paths
knitr::opts_knit$set(root.dir = ".", aliases = c(h = "fig.height", w = "fig.width", ow = "out.width"))
knitr::opts_chunk$set(
  dev = c("png"),
  fig.path = here("results", "figures", glue("{subdir_name}/")), dpi = 300
)
theme_set(theme_light()) # global option for ggplot2
```

```{r knit, echo=F, eval=FALSE}
rmarkdown::render(
  input = here("scripts", str_c(subdir_name, ".Rmd")),
  output_file = here("results", "reports", paste(subdir_name, ".html", sep="_"))
)
```

# Load data

Change filenames here to have the most updated versions if needed
These files are outputted from scripts for organising the metadata (specifically demographics_master; not uploaded to github currently) 

```{r load_data}

all_samples_demo <- read.csv(here("processed_data", "demographics_microbiome_samples_250_01092022.csv"), header=T)

demo <- read.csv(here("processed_data", "demographics_microbiome_subjects_206_01092022.csv"), header=T)

```

# Exclude those who died

```{r filter}

demo <- demo %>% subset(! (cos_excluded_options %in% c(3,4)))


all_samples_demo <- all_samples_demo %>% subset(! (cos_excluded_options %in% c(3,4)))

#keep those that went for sequencing
all_seq <- all_samples_demo %>% subset(seq_16s==1)
all_seq_subj <- unique(all_seq$record_id)
demo_seq <- demo %>% subset(record_id %in% all_seq_subj)


#only meconium samples to compare those that had enough dna for sequencing vs those with low yield
all_mec_demo <- all_samples_demo %>% subset(Visit_number=="V1")


```

# Checking normality of the variables
Do it separately for term and preterm infants
But also look at the full cohort together because I only have 16 term infants with sequencing data (and 4 had to be excluded in the end due to low read count)
Also do it only for the infants who actually have sequencing data because this is the table I will include in paper/thesis

```{r normality}

numvars <- demo_seq %>% select(GA_birth, birthweight_g, birthweight_z_score, membr_rupt_durations_hours, days_nnu,
                               totaldays, prop_abx_totaldays, GA_discharge, prop_days_excl_any_breastmilk, prop_days_excl_formula, prop_days_mixed_feed,
                               maternal_age, bmi_at_booking)

preterms_demoseq <- demo_seq %>% subset(preterm==1) %>% select(GA_birth, birthweight_g, birthweight_z_score, membr_rupt_durations_hours, days_nnu,
                               totaldays, prop_abx_totaldays, GA_discharge, prop_days_excl_any_breastmilk, prop_days_excl_formula, prop_days_mixed_feed,
                               maternal_age, bmi_at_booking)
terms_demoseq <- demo_seq %>% subset(preterm==0) %>% select(GA_birth, birthweight_g, birthweight_z_score, membr_rupt_durations_hours, days_nnu,
                               totaldays, prop_abx_totaldays, GA_discharge, prop_days_excl_any_breastmilk, prop_days_excl_formula, prop_days_mixed_feed,
                               maternal_age, bmi_at_booking)


mvn(data = numvars, univariateTest = "SW", univariatePlot = "qqplot")
mvn(data = numvars, univariateTest = "SW", univariatePlot = "histogram")

#also look in preterms only
mvn(data = preterms_demoseq, univariateTest = "SW", univariatePlot = "qqplot")
mvn(data = preterms_demoseq, univariateTest = "SW", univariatePlot = "histogram")

#no point in doing this in terms as they will only have 16 datapoints and most continuous variables are NA as NNU specific

```

Normally distributed in full dataset: birthweight_g, maternal_age
Same in preterms only

# Make demographic/clinical data descriptive table

## For all subjects

There are warnings because term infants do not have clinical data available so cannot do statistical tests - this is fine.

``` {r demotable}

allVars <- c("GA_birth", "sex", 
             "birthweight_g", "birthweight_z_score", 
             "Vaginal_delivery", 
             "antenatal_steroids_given", "mgso_any", 
             "labour_antibiotics", "membr_rupt_durations_hours",
             "days_nnu",  
             "bronchopulmonary_dysplasia", "nec_binary",
             "sepsis_binary", "retinopathy_or_premat", 
             "AB_before72h", "AB_after72h", "totaldays", "prop_abx_totaldays",
             "GA_discharge", "BM_75",
             "prop_days_excl_any_breastmilk", "prop_days_excl_formula", "prop_days_mixed_feed", "feeding_at_discharge",
             "maternal_age", "bmi_at_booking",
             "SIMD_perinatal",
             "maternal_final_education", "paternal_final_education")

catVars <- c("sex", 
             "Vaginal_delivery",
             "antenatal_steroids_given", "mgso_any", "labour_antibiotics","BM_75",
             "bronchopulmonary_dysplasia", "nec_binary", "AB_before72h", "AB_after72h",
             "sepsis_binary", "retinopathy_or_premat",
             "feeding_at_discharge",
             "SIMD_perinatal",
             "maternal_final_education", "paternal_final_education")

nnvars <- c("GA_birth", "birthweight_z_score", "membr_rupt_durations_hours", "days_nnu",
                               "totaldays", "prop_abx_totaldays", "GA_discharge", "prop_days_excl_any_breastmilk", "prop_days_excl_formula", "prop_days_mixed_feed",
                               "bmi_at_booking")


#table for all subjects who had samples collected
descr_table <- CreateTableOne(vars=allVars, strata="preterm", data=demo, factorVars = catVars, includeNA = T, addOverall = T) #errors because we have missing values
matrix_descr_table <- print(descr_table, showAllLevels = T, nonnormal = nnvars, minMax = T, explain=T, exact = catVars, contDigits = 3, printToggle = F, noSpaces = T)


#table for all subjects whose samples went to miseq sequencing
descr_table2 <- CreateTableOne(vars=allVars, strata="preterm", data=demo_seq, factorVars = catVars, includeNA = T, addOverall = T) #errors because we have missing values
matrix_descr_table2 <- print(descr_table2, showAllLevels = T, nonnormal = nnvars, minMax = T, explain=T, exact = catVars, contDigits = 3, printToggle = F, noSpaces = T)


write.csv(matrix_descr_table, file=here("results", "tables", "descriptive_stats_microb_subj_all.csv"))
write.csv(matrix_descr_table2, file=here("results", "tables", "descriptive_stats_microb_subj_miseq.csv"))


#also do table for subjects that are included in the very final dataset
#first remove some samples
final_sample_demo <- all_seq %>% subset(! sample_id %in% c("TW_03_8100_V1", "TW_03_8169_V1", "TW_03_8203_V1",
                                                           "TW_03_8206_V1", "TW_03_8368_V1", "TW_03_8370_V1",
                                                           "TW_03_8402_V1", "TW_03_8448_V1"))
final_subjects <- unique(final_sample_demo$record_id)
final_subjects_demo <- demo %>% subset(record_id %in% final_subjects)

descr_table3 <- CreateTableOne(vars=allVars, strata="preterm", data=final_subjects_demo, factorVars = catVars, includeNA = T, addOverall = T) #errors because we have missing values
matrix_descr_table3 <- print(descr_table3, showAllLevels = T, nonnormal = nnvars, minMax = T, explain=T, exact = catVars, contDigits = 3, printToggle = F, noSpaces = T)

write.csv(matrix_descr_table3, file=here("results", "tables", "descriptive_stats_microb_finalsubjects.csv"))

```

## For meconium samples (low vs sufficient DNA yield for sequencing)
Make a demographics table comparing babies whose mec samples had enough DNA to be sent to sequencing

``` {r demotable_meconium}

#first check normal distributions

mec_numvars <- all_mec_demo %>% select(GA_birth, postnatal_age_at_sample_d, PMA_sample_w,
             birthweight_g, birthweight_z_score, 
             days_nnu,  
             GA_discharge, 
             maternal_age, bmi_at_booking)

mvn(data = mec_numvars, univariateTest = "SW", univariatePlot = "qqplot")
mvn(data = mec_numvars, univariateTest = "SW", univariatePlot = "histogram")

#normally distributed: PMA_sample_w, maternal_age

allVars2 <- c("preterm", "GA_birth", "postnatal_age_at_sample_d", "PMA_sample_w",
             "sex", "birthweight_g", "birthweight_z_score", 
             "Vaginal_delivery", "antenatal_steroids_given", "mgso_any", 
             "labour_antibiotics", "AB_before72h", "AB_after72h",
             "days_nnu",  
             "bronchopulmonary_dysplasia", "nec_binary",
             "sepsis_binary", "retinopathy_or_premat",
             "GA_discharge", 
             "maternal_age", "bmi_at_booking", "SIMD_perinatal")

catVars2 <- c("preterm",
             "sex", 
             "Vaginal_delivery", "antenatal_steroids_given", "mgso_any", 
             "labour_antibiotics", "AB_before72h", "AB_after72h",
             "bronchopulmonary_dysplasia", "nec_binary",
             "sepsis_binary", "retinopathy_or_premat", "SIMD_perinatal")

nnvars2 <- c("GA_birth", "postnatal_age_at_sample_d",
             "birthweight_g", "birthweight_z_score", 
             "days_nnu",  
             "GA_discharge", 
             "bmi_at_booking")

#table for all subjects who had samples collected
descr_table3 <- CreateTableOne(vars=allVars2, strata="seq_16s", data=all_mec_demo, factorVars = catVars2, includeNA = T, addOverall = T) #errors because we have missing values
matrix_descr_table3 <- print(descr_table3, showAllLevels = T, nonnormal = nnvars2, minMax = T, explain=T, exact = catVars, contDigits = 3, printToggle = F, noSpaces = T)

write.csv(matrix_descr_table3, file=here("results", "tables", "descriptive_stats_meconium_subj_miseqvsnot.csv"))


```

Calculate some ages to include in the thesis

```{r discharge}

discharge <- all_samples_demo %>% subset(Visit_number=="V2")

#calculate how much difference is between GAdischarge and PMA at sample
discharge <- discharge %>% mutate(diff_GAdisch_GAsample = GA_discharge - PMA_sample_w)

#calculate the mean of this
mean(discharge$diff_GAdisch_GAsample)
median(discharge$diff_GAdisch_GAsample)
hist(discharge$diff_GAdisch_GAsample)

#some dates were recorded as later than actual discharge so will replace these with 0
discharge <- discharge %>% mutate(diff_GAdisch_GAsample2 = ifelse(diff_GAdisch_GAsample < 0, 0, diff_GAdisch_GAsample))

mean(discharge$diff_GAdisch_GAsample2)
median(discharge$diff_GAdisch_GAsample2)
min(discharge$diff_GAdisch_GAsample2)
max(discharge$diff_GAdisch_GAsample2)
hist(discharge$diff_GAdisch_GAsample2)
sd(discharge$diff_GAdisch_GAsample2)

#how many had samples collected earlier than X weeks of discharge (and the proportion)
sum(discharge$diff_GAdisch_GAsample2>2)
sum(discharge$diff_GAdisch_GAsample2>2)/106

sum(discharge$diff_GAdisch_GAsample2>3)
sum(discharge$diff_GAdisch_GAsample2>3)/106

sum(discharge$diff_GAdisch_GAsample2>4)
sum(discharge$diff_GAdisch_GAsample2>4)/106


#also calculate median and range for discharge sample ages
shapiro.test(discharge$postnatal_age_at_sample_d)
median(discharge$postnatal_age_at_sample_d)
min(discharge$postnatal_age_at_sample_d)
max(discharge$postnatal_age_at_sample_d)

shapiro.test(discharge$PMA_sample_w)
median(discharge$PMA_sample_w)
min(discharge$PMA_sample_w)
max(discharge$PMA_sample_w)


preterm_mec <- all_samples_demo %>% subset(Visit_number=="V1" & preterm==1)
hist(preterm_mec$postnatal_age_at_sample_d)
ggqqplot(preterm_mec$postnatal_age_at_sample_d)
hist(preterm_mec$GA_birth)
ggqqplot(preterm_mec$GA_birth)
cor.test(preterm_mec$postnatal_age_at_sample_d, preterm_mec$GA_birth, method="spearman")

#also calculate median and range for preterm mec
shapiro.test(preterm_mec$postnatal_age_at_sample_d)
median(preterm_mec$postnatal_age_at_sample_d)
min(preterm_mec$postnatal_age_at_sample_d)
max(preterm_mec$postnatal_age_at_sample_d)

shapiro.test(preterm_mec$PMA_sample_w)
median(preterm_mec$PMA_sample_w)
min(preterm_mec$PMA_sample_w)
max(preterm_mec$PMA_sample_w)


term_mec <- all_samples_demo %>% subset(Visit_number=="V1" & preterm==0)
hist(term_mec$postnatal_age_at_sample_d)
ggqqplot(term_mec$postnatal_age_at_sample_d)
hist(term_mec$GA_birth)
ggqqplot(term_mec$GA_birth)
cor.test(term_mec$postnatal_age_at_sample_d, term_mec$GA_birth, method="spearman")

#also calculate median and range for preterm mec
shapiro.test(term_mec$postnatal_age_at_sample_d)
median(term_mec$postnatal_age_at_sample_d)
min(term_mec$postnatal_age_at_sample_d)
max(term_mec$postnatal_age_at_sample_d)

shapiro.test(term_mec$PMA_sample_w)
median(term_mec$PMA_sample_w)
min(term_mec$PMA_sample_w)
max(term_mec$PMA_sample_w)

#is postnatal age different between term and preterm meconium samples
wilcox.test(term_mec$postnatal_age_at_sample_d, preterm_mec$postnatal_age_at_sample_d, paired=F) 


#how many overlapping samples obtained
length(all_samples_demo$record_id[duplicated(all_samples_demo$record_id)])

#how many in the very final dataset
length(final_sample_demo$record_id[duplicated(final_sample_demo$record_id)])

```