---
title: "`r subdir_name`"
author: "Kadi Vaher"
date: "`r Sys.time()`"
output: html_document
---

```{r setup, include=T, message=F}
subdir_name <- "microbiota_mri_demo"

# load packages
library(tidyverse)
library(here)
library(magrittr)
library(glue)
library(ggpubr)
library(knitr)
library(readxl)
library(MVN)
library(tableone)

# set paths
knitr::opts_knit$set(root.dir = ".", aliases = c(h = "fig.height", w = "fig.width", ow = "out.width"))
knitr::opts_chunk$set(
  dev = c("png"),
  fig.path = here("results", "figures", glue("{subdir_name}/")), dpi = 300
)
theme_set(theme_light()) # global option for ggplot2
```

```{r knit, echo=F, eval=FALSE}
rmarkdown::render(
  input = here("scripts", str_c(subdir_name, ".Rmd")),
  output_file = here("results", "reports", paste(subdir_name, ".html", sep="_"))
)
```

# Read in data

Subject list (those with pre-discharge samples)

```{r subj_list}

subj_list <- read.csv(here("processed_data", "disch_sample_subjects.csv"))

```

MRI-specific demographics (including parts of the MRI report)


```{r mri_demo}

mridemo <- read.csv(here("raw_data", "redcap_extractions", "EdinburghBirthCohort-KadiMRIdemo_DATA_2022-09-02_1428.csv"))

```

Tables with MRI data in order to find the max sample size that have any kind of imaging

```{r mri_data}

str_data <- read.csv(here("imaging_data", "raw_data", "Morphology-MASTER.csv"))
diff_gm <- read_excel(here("imaging_data", "raw_data", "dMRI_metrics_TEBC.xlsx"), sheet = 1)
tracts <- read.csv(here("imaging_data", "processed_data", "weighted_tracts", "FA_tracts.csv"))


```

# Wrangle data

subset the imaging datafiles to those subjects I can use (have pre-discharge samples)

```{r wrangle}

#first remove one subject from the gm diffusion file
diff_gm <- diff_gm %>% subset(!record_id=="8382")


str_data <- str_data %>% rename(record_id=subject.ID) %>%
  subset(record_id %in% subj_list$record_id)
diff_gm <- diff_gm %>% subset(record_id %in% subj_list$record_id)
tracts <- tracts %>% subset(record_id %in% subj_list$record_id)


#which babies are in tracts list but not in structural data -> need to check whether these three participants need to be excluded due to biological reasons that lead to technical difficulties (e.g. large ventricles)
tracts$record_id[which(! (tracts$record_id %in% str_data$record_id))]


#which babies are in structural list but not in tracts data -> for these we don't have diffusion data or the processing failed
str_data$record_id[which(! (str_data$record_id %in% tracts$record_id))]

#which babies are in structural list but not in gm diffusion data -> for these we don't have diffusion data or the processing failed
str_data$record_id[which(! (str_data$record_id %in% diff_gm$record_id))]

#which babies in tracts data but not in gm diffusion
tracts$record_id[which(! (tracts$record_id %in% diff_gm$record_id))]

#make a list with subjects that have any imaging data

all_imaging_subj <- c(str_data$record_id, tracts$record_id, diff_gm$record_id)
all_imaging_subj <- unique(all_imaging_subj)


#subset the demographics table to this list
demo <- mridemo %>% subset(record_id %in% all_imaging_subj)


```

# Look at the MRI reports

```{r mri_reports}

#periventricular loss score=3
demo$record_id[which(demo$mri_wm_perivent_loss==3)]
#periventricular loss score=2
demo$record_id[which(demo$mri_wm_perivent_loss==2)]

#cyst score=3
demo$record_id[which(demo$mri_wm_cystic_abnormality==3)]
#cyst score=2
demo$record_id[which(demo$mri_wm_cystic_abnormality==2)]

#ventricular dilatation score=3
demo$record_id[which(demo$mri_wm_ventric_dilatation==3)]
#ventricular dilatation score=2
demo$record_id[which(demo$mri_wm_ventric_dilatation==2)]


```
Notes about subjects to exclude
- 8342 and 8250 -> structural processing didn't work. registration done via FA so if have diffusion then have tracts (don't need structural)
- 8319: very big ventricles -> has diffusion, structural processing fails because of the large ventricles -> can keep it as the tracts we checked look fine!! 
- 8380: has both structural and the tracts -> exclude this baby too; probably a very specific injury.
- 8345: very large ventricles; have all the data there. -> can include; tracts look similar to 8319
- Parenchemal findings that mean exclusion: 8160, 8405, 8428 -> I only have 8405 to consider

# Final list of subjects

```{r final_list}

nrow(demo)

demo <- demo %>% subset(! record_id %in% c(8380, 8405, 8160, 8428))

nrow(demo)

mri_mb_subjects <- demo$record_id

#infants with structural data
str_data <- str_data %>% subset(record_id %in% mri_mb_subjects)
nrow(str_data)

#infants with diffusion data in regions (GM and WM labels)
diff_gm <- diff_gm %>% subset(record_id %in% mri_mb_subjects)
nrow(diff_gm)

#infants with tracts data
tracts <- tracts %>% subset(record_id %in% mri_mb_subjects)
nrow(tracts)

```
# Demographics/clinical data file

Now make a demographics file for all infants I will include in the microbiota-brain project

```{r read_demo}

demodata <- read.csv(here("processed_data", "demographics_microbiome_subjects_206_01092022.csv"))

all_samples_demo <- read.csv(here("processed_data", "demographics_microbiome_samples_250_01092022.csv"), header=T)

#subset to only have the pre-discharge samples
all_disch <- all_samples_demo %>% subset(Visit_number=="V2")


#read the QC data
qcdata <- read.csv(here("processed_data", "metadata_qc_decontam_23082022.csv"))
ids <- qcdata %>% select(sample_id, sample_work_nr)

all_disch <- merge(ids, all_disch, by="sample_id", all.y=T)

```

## Wrangle 
Remove some columns from the MRI demo that are already present in the big demographics file and merge

```{r demo_wrangle}

demo <- demo %>% select(record_id, mri_wm_perivent_loss, mri_wm_cystic_abnormality, 
                        mri_wm_ventric_dilatation, date_of_mri )

demo <- merge(all_disch, demo, by="record_id", all.y=T)

#remove duplicate samples
demo <- demo %>% subset(! (sample_id %in% c("TW_03_8447_V1_s2", "TW_03_8034_V2_s1", "TW_03_8049_V2_s1", "TW_03_8070_V2_s2")))

#correct the GA at mri column

demo <- demo %>% mutate(GA_mri=cga_mri+cga_mri_plus/7)

#calculate time between sample collection and MRI

demo <- demo %>% mutate(diff_predisch_mri_w = GA_mri-PMA_sample_w)

mean(demo$diff_predisch_mri_w)

#recode the maternal education column and also the paternal

demo <- demo %>% mutate(
  maternal_edu_recoded=case_when(maternal_final_education == 1 ~ "None",
                                 maternal_final_education == 2 ~ "None",
                                 maternal_final_education == 3 ~ "None",
                                 maternal_final_education == 4 ~ "High_school",
                                 maternal_final_education == 5 ~ "College/University",
                                 maternal_final_education == 6 ~ "College/University",
                                 maternal_final_education == 7 ~ "College/University"))

demo <- demo %>% mutate(
  paternal_edu_recoded=case_when(paternal_final_education == 1 ~ "None",
                                 paternal_final_education == 2 ~ "None",
                                 paternal_final_education == 3 ~ "None",
                                 paternal_final_education == 4 ~ "High_school",
                                 paternal_final_education == 5 ~ "College/University",
                                 paternal_final_education == 6 ~ "College/University",
                                 paternal_final_education == 7 ~ "College/University"))


#write into a file
#write.csv(demo, here("processed_data", "demographics_microbiome_mri_79_10102022.csv"), row.names=F)


```

# Demographics table

### Checking normality of the variables

```{r normality}

numvars <- demo %>% select(GA_birth, birthweight_g, birthweight_z_score, membr_rupt_durations_hours, days_nnu,
                               totaldays, prop_abx_totaldays, GA_discharge, prop_days_excl_any_breastmilk, prop_days_excl_formula, prop_days_mixed_feed,
                               maternal_age, bmi_at_booking, 
                           GA_mri, mri_weight, mri_weight_z_score, diff_predisch_mri_w,
                           postnatal_age_at_sample_d, PMA_sample_w)


mvn(data = numvars, univariateTest = "SW", univariatePlot = "qqplot")
mvn(data = numvars, univariateTest = "SW", univariatePlot = "histogram")

```

Normally distributed in full dataset: birthweight_g, maternal_age, GA_mri, mri_weight_z_score, diff_predisch_mri_w

## Make the descriptive table

``` {r demotable}

allVars <- c("GA_birth", 
             "sex", 
             "birthweight_g", "birthweight_z_score", 
             "GA_mri", "mri_weight", "mri_weight_z_score",
             "postnatal_age_at_sample_d", "PMA_sample_w",
             "diff_predisch_mri_w",
             "Vaginal_delivery", 
             "antenatal_steroids_given", "mgso_any", 
             "labour_antibiotics", "membr_rupt_durations_hours",
             "days_nnu",  
             "bronchopulmonary_dysplasia", "nec_binary",
             "sepsis_binary", "retinopathy_or_premat", 
             "AB_before72h", "AB_after72h", "totaldays", "prop_abx_totaldays",
             "GA_discharge", 
             "prop_days_excl_any_breastmilk", "prop_days_excl_formula", "prop_days_mixed_feed", "feeding_at_discharge",
             "mri_current_feeding",
             "maternal_age", "bmi_at_booking",
             "SIMD_perinatal",
             "maternal_final_education", "paternal_final_education",
             "maternal_edu_recoded", "paternal_edu_recoded")

catVars <- c("sex", 
             "Vaginal_delivery",
             "antenatal_steroids_given", "mgso_any", "labour_antibiotics",
             "bronchopulmonary_dysplasia", "nec_binary", "AB_before72h", "AB_after72h",
             "sepsis_binary", "retinopathy_or_premat",
             "feeding_at_discharge","mri_current_feeding",
             "SIMD_perinatal",
             "maternal_final_education", "paternal_final_education",
             "maternal_edu_recoded", "paternal_edu_recoded")

nnvars <- c("GA_birth", "birthweight_z_score", "membr_rupt_durations_hours", "days_nnu",
                               "totaldays", "prop_abx_totaldays", "GA_discharge", "prop_days_excl_any_breastmilk", "prop_days_excl_formula", "prop_days_mixed_feed",
                               "bmi_at_booking", "mri_weight", "postnatal_age_at_sample_d", "PMA_sample_w")


#table for all subjects who had samples collected
descr_table <- CreateTableOne(vars=allVars, data=demo, factorVars = catVars, includeNA = T, addOverall = T)
matrix_descr_table <- print(descr_table, showAllLevels = T, nonnormal = nnvars, minMax = T, explain=T, exact = catVars, contDigits = 3, printToggle = F, noSpaces = T)

write.csv(matrix_descr_table, file=here("results", "tables", "descriptive_stats_microb_match_mri_79.csv"))

```